---
title: "Week 3 Assignment"
author: Steven Spiegel
output: pdf_document
date: "2023-09-05"
---


### Assignment 3 responses 

Below are my responses for Assignment 3.  

**Create a new column in the AmesData data frame which has a value of 1 if the house is built in year 2000 or later or 0 if it is built before year 2000 using the following method.**

Question 1: **ifelse(name this column – w3ifelse).  What is the first 10 elements of w3ifelse?**

Response: **See code below**

```{r Q1}
library(AmesHousing)
library(dplyr)
# Load data
AmesData <- make_ames()

AmesData$w3ifelse <- if_else(AmesData$Year_Built >= 2000, 1, 0)
head(AmesData$w3ifelse, 10)
```

Question 2: **if-Else and a For loop (name this column – w3for)  What is the first 10 elements of w3for?**

Response: **See code below**

```{r Q2}
AmesData$w3for = numeric(nrow(AmesData)) # instantiate column with zeros
for (i in 1:nrow(AmesData)) {
  
  if (is.na(AmesData$Year_Built[i])){
    AmesData$w3for[i] <- NA
  } else if(AmesData$Year_Built[i] >= 2000){
    AmesData$w3for[i] <- 1
  } else {
    AmesData$w3for[i] <- 0 
  }
}

head(AmesData$w3for, 10)
```


Question 3: **Build your own function and use sapply (name this column - w3apply)  What is the first 10 elements of w3apply?**

Response: **See code below**

```{r Q3}

past2000 <- function(x) {
  if (is.na(x)) return (x)
  else if ( x >=2000) return (1)
  else return (0)
}

AmesData$w3apply <- sapply(AmesData$Year_Built, past2000)
head(AmesData$w3apply, 10)
```

Question 4: **Create a column w3diff which is the difference of the two columns w3ifelse and w3apply. DO NOT use a for loop or the apply set of functions for this step. Simple subtraction will work. What is the total sum of this column?**


Response: **See code below**

```{r Q4}
AmesData$w3diff <- AmesData$w3apply - AmesData$w3for
sum(AmesData$w3diff)

```

As we can see, the difference is 0.  



